---
title: "Joshua Froess: Homework C for 431"
author: "Joshua Froess"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    code_folding: show
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(comment=NA)
options(width = 70)
library(MASS); library(e1071); library(tidyverse)

```

```{r create_hwC_tibble}
hwC <- tbl_df(faithful)
hwC
```

# Question 1

The histogram below shows the average waiting times until the eruptions of the Old Faithful Geyser.  The wait times seem to peak between 75 to 85 minutes with most of them being around 80 minutes long. There is another peak around 55 minutes showing that this histogram is bimodal.

```{r q01_histogram}
ggplot(hwC, aes(x = waiting)) +
  geom_histogram(binwidth = 3,
  fill = "firebrick3", col = "white") +
  labs(x = "Waiting Time (in minutes)",
       y = "Number of Eruptions",
       title = "Time Between Eruptions for the Old Faithful Geyser") +
  theme_bw()
       
```

# Question 2

The typical waiting time is about 71 minutes when looking at the distrubition among the mean, median, and 80% mean. The median is slightly higher then this but this could also be because the peak distrubition on the histogram is around 80 minutes.

```{r q02_trimmed_mean_durations}
hwC %>%
  summarise(Mean = mean(waiting),
            Median = median(waiting),
            Mean80 = mean(hwC$waiting, trim = 0.1),
            Stan_Dev = sd(hwC$waiting),
            IQR = IQR(hwC$waiting))

```

# Question 3

```{r}
summary(hwC$waiting)
```


The interquartile range is 24, and goes from 58 to 82 minutes. The interquartile range represents the middle 50% of the waiting times. The standard deviation on the other hand is 13.6, which means + or - 13.6 from the mean will represent 68% of the data. These are two different measure to show where majority of the data is located.

# Question 4

The distribution on the graph looks to be multimodal. There seems to be two peaks among the histogram one at 55 minutes and one at 80 minutes. The first distribution on the graph is slightly smaller then the second distrubition.

# Question 5

```{r q95_skewness}
skewness(hwC$eruptions)
```


The distribution is skewed to the left because the skewness value shown is negative. This is also because the median is higher then the mean. The histogram also shows that the distribution is skewed to the left. The first modal distribution in the graph is lowering the mean since more values are presented in the second modal distribution that is farther to the right.

# Question 6

There looks to be very few eruptions that had a waiting times around 95 minutes. There are also few eruptions between the two distributions. Both of these could be outliers for the waiting time until the next eruption occurs.

# Question 7

No because the data that was collected about the waiting times between eruptions was not a normal distribution.

# Question 8

This scatterplot shows the relationship between the duration and time until the next eruption. A linear line was used because the correlation between these two values looks to be strong linear relationship.

```{r q08_scatterplot_draft}
ggplot(hwC, aes(x = eruptions, y = waiting)) +
  geom_point(size = 1.5, col = "black") +
  geom_smooth(method = "lm", col = "navyblue") +
  labs(x = "Duration of Eruptions (mins)",
       y = "Time Between Eruptions (mins)",
       title = "Old Faithful Geyser Eruptions",
       subtitle = "Comparing the Duration to the Time Between Eruptions") +
  theme_bw()

```


# Question 9

The correlation between the duration of eruptions and the waiting time between eruptions is about 0.9. This means there is a strong linear positive correlation between these measured times of the old faithful geyser eruption.

```{r q09_correlation}
cor(hwC$waiting, hwC$eruptions)
```


# Question 10

Yes, a linear model would be appropriate to use. The scatter plot shows a strong linear relationship between the duration of eruptions and waiting time between eruptions. The correlation is also close to +1 which means the linear relationship between these two variables is strong.

# Question 11

```{r q11_create_hwC}
hwCextra <- tbl_df(MASS::geyser)
hwCextra
```


## 11a. Question 8 for the new data

This scatterplot shows the relationship between the duration and the time until that eruption. This graph uses the geyser dataset, which redescribes the waiting variable and has 299 observations instead of 272 observations. The relationship between waiting and eruptions is slightly weaker in this graph with the dots being a little more spread out. This still shows a moderate correlation between waiting and eruptions. 

```{r q11a_scatterplot_geyser}
ggplot(hwC, aes(x = eruptions, y = waiting)) +
  geom_point(size = 1.5, col = "black") +
  geom_smooth(method = "lm", col = "navyblue") +
  labs(x = "Duration of Eruptions (mins)",
       y = "Time Until Eruption Happens (mins)",
       title = "Old Faithful Geyser Eruptions",
       subtitle = "Comparing the Duration to the Time Until the Eruption   Happens") +
  theme_bw()

```


## 11b. Question 9 for the new data

```{r q11b_correlation}
cor(hwCextra$duration, hwCextra$waiting)

```

The correlation of the geyser set is around -0.64. This means the two variables are moderately correlated in a negative linear relationship. This correlation is slightly less strong because the waiting variable is defined in a different way among the geyser set. Waiting is now the time until the eruption happens and eruption is the duration of the eruption. There is also now 299 observations and previously there was only 272 observations. All of these factors could attribute from this becoming a moderate correlation instead of a strong correlation.

## 11c. Question 10 for the new data

A linear model would still be appropriate even though the correlation is slightly lower. A lot of the dots cluster around the linear line. The correlation is not as strong as previously but is still good enough to be used to show the relationship between duration and time until the eruption.

# Session Information


```{r}
sessionInfo()
```