---
title: "Joshua Froess: Homework 5 for 435"
author: "Joshua Froess"
date: "2/26/2020"
output:
  html_document:
    toc: yes
    code_folding: show
---

## Initial Setup and Loading Libraries

```{r setup, message=FALSE}
knitr::opts_chunk$set(comment = NA)
options(width = 70)
```

```{r, message=FALSE}
library(magrittr); library(janitor); library(here)
library(survival); library(patchwork); library(survminer)
library(tidyverse)
```

# Loading Data

First the data about patients with a malignant tumour in the kidney is loaded.

```{r, message=FALSE}
kidney <- read_csv(here("data", "kidney.csv"))

head(kidney)
```

Next a new variable is created called `km_cat`. This variable will be used to create a kaplan-meier graph later on for the specific groups 1 and 2 in the `km_cat` variable. A sanity check is also done to make sure the variable was properly made and filtered down to only the categories 1 and 2.

```{r}
kidney_km <- kidney %>%
  mutate(km_cat = ifelse(age < 2 & neph < 1, 1,
                  ifelse(age > 1 & age < 3 & neph > 0, 2, 3))) %>%
  filter(km_cat == 1 | km_cat == 2)

kidney_km %>%
  tabyl(km_cat)
```

# Cox Model

The cox model on the survival of the patients with kidney tumours shows that the nephrectomy status of patients is statistically detectable at the 95% level. This means that patients that got a nephrectomy are statistically different then patients that did not get a nephrectomy. It should be kept in mind that the sample size for patients that did not get a nephrectomy is realitvely small and may not be completely representative of this procedure.

```{r}
coxph(Surv(survt, censor) ~ age + neph, data = kidney)
```

# Kaplan-Meier Plot

A kaplan-meier plot is then made for patients by nephrectomy status and age. Specifically, the red line represents patients that did not have a nephrectomy and were under the age of 60. The blue line represents patients that had a nephrectomy and were between the ages of 60 and 70.

```{r}
ggsurvplot(
  fit = survfit(Surv(survt, censor) ~ km_cat, data = kidney_km),
  xlab = "Days",
  ylab = "Survival Probability",
  title = "Patients with kidney tumour by nephrectomy and age status",
  legend = "bottom",
  legend.labs = c("no nephrectomy & age < 60", "nephrectomy & 60 < age < 70"),
  pval = TRUE)
```

