---
title: 'Homework 1: PQHS 500'
author: "Joshua Froess"
date: "1/22/2020"
output: html_document
---

# Initial Setup Loading Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(magrittr); library(janitor); library(here)
library(simputation); library(tidyverse)
```

# Question 2: Mock Proprosal

Cardiovascular disease is a leading cause of death in the United States and myocardial infarctions are one aspect of this. Using the digitalis investigation group database the effects of the percentage of ejection fraction on myocardial infarction can be studied. Low rates of ejection fraction percent can possibly be an indicator of heart failure. Specifically, this can study if low percentages of ejection fraction are correlated with higher rate of myocardial infarction in adults. This is important because further understanding indicators of specific cardiovascular disease can lead to faster treatment. Possible early treatment to prevent myocardial infarctions are important to lowering rates of cardiovascular disease death.

The exposure will be categorical because the percentage of ejection fraction is usually put into three groups: low, normal, and high. The outcome will be binary on whether a patient had a myocardial infarction. There are many possible covariates that will need adjusted for such as the treatment of placebo or digoxin. Whether the patient was on the placebo or digoxin could affect the outcome of being hospitalized by a myocardial infarction. Other cardiovascular levels will need to be adjusted for as well to completely understand the correlation of ejection fraction and myocardial infarction. These other covariates will be: history of hypertension, bmi, blood pressurs (systolic and diastolic), using hydralazine, nitrates, history of diabetes, and current angina. All of these covariates will be kept constant during the analysis to best understand how ejection fraction is correlated to myocardial infarction.

# Question 3

## Loading Data

```{r}
dig <- read.csv(here("data", "dig.csv")) %>%
  clean_names()
```

## Tidy Data

```{r}
dig_tidy <- dig %>%
  select(id, trtmt, death, age, race, sex, ejf_per, bmi, functcls, angina)

colSums(is.na(dig_tidy))
```

## Imputation

```{r}
set.seed(20202201)

dig_imputed <- dig_tidy %>%
  mutate(bmi = bmi,
         functcls = functcls,
         angina = angina) %>%
  data.frame() %>%
  impute_rhd(., bmi ~ 1) %>%
  impute_rhd(., functcls ~ 1) %>%
  impute_rhd(., angina ~ 1) %>%
  tbl_df()

colSums(is.na(dig_imputed))
```

## Creating training and test samples

```{r}
set.seed(20202201)

dig_training <- dig_imputed %>% sample_n(5000)
dig_test <- anti_join(dig_imputed, dig_training, by = "id")

dim(dig_imputed)
dim(dig_training)
dim(dig_test)
```

## Logistic Regression

For both the training sample and the test sample the ejection fraction percentage and NYHA functional class are statistically significant results for predicting death. The treatment of digoxin does not seem to be significant for predicting death. This is important because the main point of the DIG study was to see the effects of digoxin on death of cardiovascular disease.

```{r}
dig_logistic <- glm(death ~ trtmt + age + race + sex + ejf_per + bmi + functcls + angina,
                    family = binomial(), data = dig_training)

summary(dig_logistic)
```

```{r}
dig_logistic2 <- glm(death ~ trtmt + age + race + sex + ejf_per + bmi + functcls + angina,
                    family = binomial(), data = dig_test)

summary(dig_logistic2)
```

# Session Information

```{r}
sessionInfo()
```

