---
title: "Joshua Froess: Homework 2 for 500"
author: "Joshua Froess"
date: "1/30/2020"
output:
  html_document:
    toc: yes
    code_folding: show
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(comment = NA)
options(width = 70)
```

## Loading Packages and Data

```{r message=FALSE}
library(magrittr); library(janitor); library(here)
library(tableone); library(broom); library(tidyverse)
```

```{r}
dig <- read.csv(here("data", "dig.csv")) %>%
  clean_names()
```

# Question 1

The data was set to a random sample of 1000 observations and no missing data appeared for the previous Myocardial Infraction category.

```{r}
set.seed(20203001)

dig <- dig %>%
  sample_n(1000)

dig %>%
  tabyl(prevmi)
```

# Question 2

The data was then selected down to the outcome variable, the 12 predictor variables, and the id for the patients. Next, categorical variables were mutated so that the categories were more descriptive. Variables were also renamed to make more sense then they previously did. The data was also checked for missing data and 3 variables have missing data.

```{r}
dig1 <- dig %>%
  select(id, prevmi, trtmt, age, race, sex, ejf_per,
         chestx, bmi, klevel, creat, chfdur, exertdys,
         functcls) %>%
  mutate(race = fct_recode(factor(race),
                           White = "1",
                           Non_White = "2"),
         sex = fct_recode(factor(sex),
                          Male = "1",
                          Female = "2"),
         exertdys = fct_recode(factor(exertdys),
                               Non_Unknown = "0",
                               Present = "1",
                               Past = "2",
                               Present_and_past = "3"),
         functcls = fct_recode(factor(functcls),
                               I = "1",
                               II = "2",
                               III = "3",
                               IV = "4")) %>%
  rename(dyspnea_exert = exertdys,
         nyha_funct_class = functcls,
         chest_xray = chestx,
         chf_dur = chfdur,
         creat_lvl = creat,
         k_lvl = klevel)
colSums(is.na(dig1))
```

Multiple Normal Q-Q plots were made for the quantitative variables. This was done to assess the normality of the variables to determine if mean or median should be used. Many of the variables are seen to have outliers or skew. This will be taken into account when building the table one.

```{r message=FALSE}
p1 <- ggplot(dig1, aes(sample = age)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "Age")

p2 <- ggplot(dig1, aes(sample = bmi)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "BMI")

p3 <- ggplot(dig1, aes(sample = ejf_per)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "Ejection Fraction %")

p4 <- ggplot(dig1, aes(sample = chest_xray)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "Chest X-Ray")

p5 <- ggplot(dig1, aes(sample = k_lvl)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "Potassium Level")

p6 <- ggplot(dig1, aes(sample = creat_lvl)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "Creatine Level")

p7 <- ggplot(dig1, aes(sample = chf_dur)) +
  geom_qq() + geom_qq_line(col = "firebrick4") +
  labs(title = "CHF Duration")

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, nrow = 2,
                        top = "Normal QQ Plots of Continuous Variables")
```

The table below looks at 12 variables by whether a patient previously had a myocardial infraction. Only three variables have a statistically detectable difference at the 10% level. A higher percentage of females have previously had a myocardial infraction. Overall, there is more patients with a previous myocardial infraction the without. There is also a total of 134 patients with a missing variable, this could be affecting some of the results with missing values. Specifically, potassium levels have 130 missing values so this could be highly affected from missing data.

```{r}
dig.vars <- c("trtmt", "age", "race", "sex", "ejf_per",
              "chest_xray", "bmi", "k_lvl", "creat_lvl", "chf_dur",
              "dyspnea_exert", "nyha_funct_class")

dig.trt <- c("prevmi")

dig.tbl1 <- CreateTableOne(data = dig1,
                           vars = dig.vars,
                           strata = dig.trt)

print(dig.tbl1,
      nonnormal = c("bmi", "ejf_per", "k_lvl",
                    "creat_lvl", "chf_dur", "chest_xray"))

```

# Question 3

A logistic regression was ran on the outcome variable with the 12 predictors at a 90% confidence level. Based on the confidence intervals that the model produced three variables are statistically detectable at the 10% level. Based on this sample size it a difference is not seen with the other variables, but this could change because of the missing variables.

```{r}
m1 <- glm(prevmi ~ trtmt + age + race + sex + ejf_per +
            chest_xray + bmi + k_lvl + creat_lvl + chf_dur +
            dyspnea_exert + nyha_funct_class, family = "binomial",
          data = dig1)
tidy(m1, conf.int = TRUE, conf.level = 0.90) %>%
  select(term, estimate, std.error, conf.low, conf.high) %>%
  knitr::kable(digits = 3)
```

The logistic regression has a 134 total subjects that are missing a variable. This could affect the logistic regression. Most of the missing variables are among the potassium serum level.

```{r}
dig1 %>%
  filter(complete.cases(k_lvl, chf_dur, nyha_funct_class)) %$%
  mosaic::favstats(id ~ 1)
```

# Question 4

The data was filtered to no longer include missing data and is now limited to 866 observations instead of the original 1000.

```{r}
dig2 <- dig1 %>%
  filter(complete.cases(k_lvl, chf_dur, nyha_funct_class))

colSums(is.na(dig2))
```

The table one was remade using the new data with no missing data and all three variables stayed statistically detectable at the 10% level. All the variables look to have changed slightly now that all the missing data had been filtered out. This is to be expected with making the sample size smaller then the original 1000 observations.

```{r}
dig.vars <- c("trtmt", "age", "race", "sex", "ejf_per",
              "chest_xray", "bmi", "k_lvl", "creat_lvl", "chf_dur",
              "dyspnea_exert", "nyha_funct_class")

dig.trt <- c("prevmi")

dig.tbl1 <- CreateTableOne(data = dig2,
                           vars = dig.vars,
                           strata = dig.trt)

print(dig.tbl1,
      nonnormal = c("bmi", "ejf_per", "k_lvl",
                    "creat_lvl", "chf_dur", "chest_xray"))
```

The logistic regression model with the missing variables filtered out still has the same statistically detectable variables. The confidence intervals overall change slightly for all the variables. This is to be expected with lowering the sample size.

```{r}
m2 <- glm(prevmi ~ trtmt + age + race + sex + ejf_per +
            chest_xray + bmi + k_lvl + creat_lvl + chf_dur +
            dyspnea_exert + nyha_funct_class, family = "binomial",
          data = dig2)
tidy(m2, conf.int = TRUE, conf.level = 0.90) %>%
  select(term, estimate, std.error, conf.low, conf.high) %>%
  knitr::kable(digits = 3)
```

# Question 5

The predicted value error is rounded to the nearest 2 digits and the previous myocardial infarction was mutated to a yes and no.

```{r}
m2_aug <- augment(m2) %>%
  mutate(.fitted = round(.fitted, digits = 2),
         prevmi = fct_recode(factor(prevmi),
                             Yes = "1",
                             No = "0"))
```

The predicted variable error is plotted against the observed variables for the patients who previously had a myocardial infarction and those that did not. Majority of the prediction errors seem to be in the center for both the yes and the no categories.

```{r}
ggplot(m2_aug, aes(x = .fitted, y = prevmi, col = .fitted)) +
  geom_count() +
  labs(y = "Previous Myocardial Infarction",
       title = "Observed values vs Predicted values",
       subtitle = "for patients with a previous MI using 12 predictor variables") +
  theme_bw()
```


## Session Info

```{r}
sessionInfo()
```

